---
title: "Storm"
author: "Ishikawa"
date: "2019/7/7"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
  word_document: default
---

## The Analysis the Weather Event Impact to Population Health and Properties


### Qestions:

Across the United States,  
 1) Which types of events are most harmful with respect to population health?  
 2) Which types of events have the greatest economic consequences?  

From the Storm Events data, answer the questions.


```{r message=FALSE}
library(tidyverse)
library(lubridate)
options(digits=4)
options(warn=-1)
options(width=110)
options(scipen=3)
```

### Loading and preprocessing the data

Data from:  
 National Climatic Data Center Storm Events  
 (https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)  

First time, Read data file each variavles as character.
```{r}
Storm_wk <- read_csv('repdata_data_StormData.csv.bz2', col_names=T, col_types=str_dup('c',37))

dim(  Storm_wk) %>% cat ('Dimension: ',.,'\n')
names(Storm_wk) %>% cat ('Col Names: ',.,'\n')
```

Relational Valiables for the Questions in the file.
 EVTYPE - Event Type  
 FATALITIES  
 INJURIES  
 PROPDMG - Property Damage,  PROPDMGEXP - Exponents  
 CROPDMG - Crop Damage,      CROPDMGEXP - Exponents  

Check the Variable Data characteristics..
```{r}
map(c('EVTYPE', 'FATALITIES', 'INJURIES', 'PROPDMG', 'PROPDMGEXP', 'CROPDMG', 'CROPDMGEXP'),
    function(x){
        cat('\n',x,'---\n')
        wk1 <- Storm_wk %>% pull(x) 
        wk1 %>% is.na() %>% sum() %>% cat('Number of NAs: ',.,'\n')
        wk2 <- wk1 %>% unique()
        wk2 %>% length() %>% cat('Unique Length: ',.,'\n')
        if(length(wk2)<=20){wk2 %>% sort() %>% cat('\n')}
        else               {wk2 %>% sort() %>% head(6) %>% cat('Head: ',.,'\n')
                            wk2 %>% sort() %>% tail(6) %>% cat('Tail: ',.,'\n')}
    }) -> null
```

```{r}
Storm_wk <- Storm_wk %>% 
         select(BGN_DATE,EVTYPE,FATALITIES,INJURIES,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP) %>%
         mutate(BGN_DATE   = mdy_hms(BGN_DATE), 
                EVTYPE     = str_to_upper(EVTYPE),
                FATALITIES = as.integer(FATALITIES),
                INJURIES   = as.integer(INJURIES),
                PROPDMG    = as.double(PROPDMG),
                PROPDMGEXP = str_to_upper(PROPDMGEXP),
                CROPDMG    = as.double(CROPDMG),
                CROPDMGEXP = str_to_upper(CROPDMGEXP)) %>%
         filter(!str_detect(EVTYPE,'SUMMARY'))
dim(Storm_wk)  # [1] 902221      8
Storm_wk %>% head()
```
### Check Summary of Data
```{r}
Storm_wk %>% pull(EVTYPE) %>% unique() %>% length()
summary(Storm_wk[,c(3,4,5,7)])
Storm_wk %>% pull(PROPDMGEXP) %>% unique() %>% sort()
Storm_wk %>% pull(CROPDMGEXP) %>% unique() %>% sort()
```

### Clean Event types

There are variations in experssion in adjective for events and event names.
So, remove adjectives and be just events. and aggregate event names.

```{r}
Storm <- Storm_wk %>% 
         mutate(EVTYPE = str_replace(EVTYPE,'  ',' '),
                EVTYPE = str_replace(EVTYPE,
                         '^BITTER |^RECORD |^EXCESSIVE |^EXTREME |^SMALL |^SEVERE |^HEAVY ',''),
                EVTYPE = str_replace(EVTYPE,
                         '^HIGH |^STRONG |^PROLONG |^UNSEASONABLY |^MODERATE |^DENSE ',''),
                EVTYPE = str_replace(EVTYPE,'^BLIZZARD.*'            ,'BLIZZARD'),
                EVTYPE = str_replace(EVTYPE,'^HAIL.*'                ,'HAIL'),
                EVTYPE = str_replace(EVTYPE,'^HURRICANE.*|^TYPHOON.*','HURRICANE'),
                EVTYPE = str_replace(EVTYPE,'^LIGHTNING.*'           ,'LIGHTNING'),
                EVTYPE = str_replace(EVTYPE,'^THUNDERSTORM.*|^TSTM.*','THUNDERSTORM WIND'),
                EVTYPE = str_replace(EVTYPE,'^COASTAL.*|^BEACH.*'    ,'COASTAL FLOOD'),
                EVTYPE = str_replace(EVTYPE,'^FLASH FLOOD.*'         ,'FLASH FLOOD'),
                EVTYPE = str_replace(EVTYPE,'^FLOOD.*'               ,'FLOOD'),
                EVTYPE = str_replace(EVTYPE,'^URBAN/SML STREAM FLD'  ,'FLOOD'),
                EVTYPE = str_replace(EVTYPE,'^URBAN FLOOD'           ,'FLOOD'),
                EVTYPE = str_replace(EVTYPE,'^RIVER FLOOD.*'         ,'FLOOD'),
                EVTYPE = str_replace(EVTYPE,'^FUNNEL CLOUD.*'        ,'FUNNEL CLOUD'),
                EVTYPE = str_replace(EVTYPE,'^COLD.*'                ,'COLD/WIND CHILL'),
                EVTYPE = str_replace(EVTYPE,'^UN.+ COLD'             ,'COLD/WIND CHILL'),
                EVTYPE = str_replace(EVTYPE,'^HEAT.*|^WARM.*'        ,'HEAT'),
                EVTYPE = str_replace(EVTYPE,'^RAIN.*'                ,'RAIN'),
                EVTYPE = str_replace(EVTYPE,'^MARINE TSTM.*'         ,'MARINE THUNDERSTORM WIND'),
                EVTYPE = str_replace(EVTYPE,'^MARINE THUNDERSTORM.*' ,'MARINE THUNDERSTORM WIND'),
                EVTYPE = str_replace(EVTYPE,'^SNOW.*'                ,'SNOW'),
                EVTYPE = str_replace(EVTYPE,'^STORM SURGE.*'         ,'STORM SURGE/TIDE'),
                EVTYPE = str_replace(EVTYPE,'^SURF.*'                ,'SURF'),
                EVTYPE = str_replace(EVTYPE,'^TORNADO.*'             ,'TORNADO'),
                EVTYPE = str_replace(EVTYPE,'^WIND.*'                ,'WIND'),
                EVTYPE = str_replace(EVTYPE,'^WATERSPOUT.*'          ,'WATERSPOUT'),
                EVTYPE = str_replace(EVTYPE,'^WILD/FOREST FIRE.*'    ,'WILDFIRE'),
                EVTYPE = str_replace(EVTYPE,'^WINTER WEATHER.*'      ,'WINTER WEATHER')
        )
```

### Data Processing for Q1

Which types of events are most harmful with respect to population health?  

1) summarize sum of FATALITIES and INJURIES by each event.
2) Sort them in descending order of sum of FATALITIES and INJURIES.
3) Select events by top 10 and make pie chart.

```{r}
event_FAIN <-Storm %>% 
        mutate(FATAL_INJU = FATALITIES+INJURIES) %>%
        group_by(EVTYPE) %>% 
        summarise(FATAL_INJU=sum(FATAL_INJU)) %>% 
        arrange(desc(FATAL_INJU))
event_FAIN <- event_FAIN %>%
        head(10) %>%
        bind_rows(tibble(EVTYPE='others', FATAL_INJU=sum(event_FAIN$FATAL_INJU[11:nrow(event_FAIN)]))) %>%
        mutate(percentage=100*FATAL_INJU/sum(FATAL_INJU))
event_FAIN %>% print(n=Inf)
event_FAIN <- event_FAIN %>%
        mutate(EVTYPE=factor(EVTYPE,levels=event_FAIN$EVTYPE))

ggplot(event_FAIN) + 
    geom_bar(aes(x=1,y=FATAL_INJU,fill=EVTYPE),stat='identity') + 
    annotate('text', x=1, y=sum(event_FAIN$FATAL_INJU)-event_FAIN$FATAL_INJU[1]/2, 
             label=sprintf('%s: %d %%',
                           event_FAIN$EVTYPE[1],
                           as.integer(100*event_FAIN$FATAL_INJU[1]/sum(event_FAIN$FATAL_INJU)))) +
    coord_polar(theta='y') + scale_y_reverse() + theme_void() +
    theme(axis.title.x=element_blank(), axis.title.y=element_blank()) + 
    ggtitle('fig1. The top 10 events; Percentage of sum of FATALITIES and INJURIES')
```

### Data Processing for Q2

Which types of events have the greatest economic consequences?  

1) Create ECO_DAMAGE as valiable from PROPDMG, CROPDMG, PROPDMGEXP and CROPDMGEXP.
Exponents means Power of 10, and "H" is 10^2, "M" is Millon, "B" is Billion, and others not numeric is 10^0.  
2) summarize ECO_DAMAGE by event.
3) Sort them in descending order of ECO_DAMAGE.
4) Select events bytop 10 ECO_DAMAGE and make pie chart.

```{r}
event_DMG <- Storm %>%
        mutate(PROPDMGEXP=if_else(is.na(PROPDMGEXP),'0',PROPDMGEXP),
               CROPDMGEXP=if_else(is.na(CROPDMGEXP),'0',CROPDMGEXP),
               PROPDMGEXP=recode(PROPDMGEXP, '+'='0','-'='0','?'='0','H'='2','K'='3','M'='6','B'='9'),
               CROPDMGEXP=recode(PROPDMGEXP, '+'='0','-'='0','?'='0','H'='2','K'='3','M'='6','B'='9'),
               ECO_DAMAGE= PROPDMG*(10^as.integer(PROPDMGEXP))+CROPDMG*(10^as.integer(CROPDMGEXP))) %>% 
        group_by(EVTYPE) %>% 
        summarise(ECO_DAMAGE=sum(ECO_DAMAGE)) %>% 
        arrange(desc(ECO_DAMAGE))
event_DMG <- event_DMG %>%
        head(10) %>%
        bind_rows(tibble(EVTYPE='others', ECO_DAMAGE=sum(event_DMG$ECO_DAMAGE[11:nrow(event_DMG)]))) %>%
        mutate(percentage=100*ECO_DAMAGE/sum(ECO_DAMAGE))
event_DMG %>% print(n=Inf)
event_DMG <- event_DMG %>%
        mutate(EVTYPE=factor(EVTYPE,levels=event_DMG$EVTYPE))

ggplot(event_DMG) + 
    geom_bar(aes(x=1,y=ECO_DAMAGE,fill=EVTYPE),stat='identity') + 
    annotate('text', x=1, y=sum(event_DMG$ECO_DAMAGE)-event_DMG$ECO_DAMAGE[1]/2, 
             label=sprintf('%s: %d %%',
                           event_DMG$EVTYPE[1],
                           as.integer(100*event_DMG$ECO_DAMAGE[1]/sum(event_DMG$ECO_DAMAGE)))) +
    coord_polar(theta='y') + scale_y_reverse() + theme_void() +
    theme(axis.title.x=element_blank(), axis.title.y=element_blank()) + 
    ggtitle('fig2. The top 10 events; Percentage of sum of Property and Crops Damages')
```


### The Results

#### For Question 1

The type of events most harmful with respect to population health is TORNADO.
They occupy about 62% of the whole.
(Refer fig1)

#### For Question 2

The types of events the greatest economic consequences is HURRICANE.  
They occupy about 74% of the whole.
(Refer fig2)






